\section{Appendix}
\label{sec:appendix}

\noindent This appendix retains only compact audit material that supports the accepted empirical and DTSE chapters. Core chapter argumentation, scenario framing, and methodology exposition remain in the main body.

\subsection{DTSE Scenario-Formula Map}
\label{app:dtse_scenario_formula_map}
Table~\ref{tab:app_dtse_scenario_formula_map} gives a compact reader-facing summary of why each scenario is included, how it enters the evaluator, and which fixed settings matter most for interpretation.

\begin{table}[H]
    \centering
    \small
    \renewcommand{\arraystretch}{1.15}
    \setlength{\tabcolsep}{4pt}
    \begin{tabularx}{\textwidth}{@{}>{\raggedright\arraybackslash}p{0.18\textwidth}>{\raggedright\arraybackslash}p{0.20\textwidth}>{\raggedright\arraybackslash}p{0.27\textwidth}X@{}}
        \toprule
        \textbf{Scenario} & \textbf{Why Included} & \textbf{Rule Summary} & \textbf{Key Fixed Settings} \\
        \midrule
        Baseline Neutral &
        Establish the reference path for timing and magnitude comparisons. &
        Stable demand and background settings generate the neutral baseline against which all later deviations are measured. &
        No adverse event trigger; no signal threshold needs to be crossed for the baseline to exist as a reference path. \\
        \addlinespace
        Demand Contraction &
        Test reward--demand decoupling under weakening service use. &
        Demand decays through the scenario-specific contraction path under higher volatility. &
        Utilization becomes a material signal once it departs ten percent from the baseline path. \\
        \addlinespace
        Liquidity Shock &
        Test fast transmission from token-liquidity stress into provider-economics fields. &
        A discrete unlock-style sell event is introduced under limited market depth, producing a sudden price dislocation. &
        Event week: 20. Sell-pressure scalar: 35\%. Material price deviation threshold: 10\%. \\
        \addlinespace
        Competitive Yield Pressure &
        Isolate provider reallocation pressure without requiring demand collapse. &
        External yield pressure raises switching sensitivity, with provider types responding at different intensities under the same outside opportunity. &
        External-yield scalar: 1.5$\times$. Provider-count threshold: 5\%. Churn-increase threshold: 20\%. \\
        \addlinespace
        Provider Cost Inflation &
        Test Subsidy-Gap exposure under higher operating costs. &
        Provider cost baselines are shifted upward across the active provider set while all other scenario channels remain neutral. &
        Cost multiplier: 1.35$\times$. Profitability threshold: 10\%. Provider-count threshold: 5\%. \\
        \bottomrule
    \end{tabularx}
    \caption{Scenario logic and key fixed settings for the DTSE experiment set.}
    \label{tab:app_dtse_scenario_formula_map}
\end{table}

\subsection{DTSE Threshold Definitions}
\label{app:dtse_threshold_table}
Table~\ref{tab:app_dtse_threshold_registry_excerpt} summarizes the main fixed thresholds referenced in Sections~\ref{sec:dtse_methodology} and~\ref{sec:simulation_results}.

\begin{table}[H]
    \centering
    \footnotesize
    \renewcommand{\arraystretch}{1.15}
    \setlength{\tabcolsep}{4pt}
    \begin{tabularx}{\textwidth}{@{}>{\raggedright\arraybackslash}p{0.27\textwidth}>{\raggedright\arraybackslash}p{0.11\textwidth}>{\raggedright\arraybackslash}p{0.12\textwidth}X@{}}
        \toprule
        \textbf{Threshold or Setting} & \textbf{Value} & \textbf{Unit} & \textbf{Interpretive Role} \\
        \midrule
        Negative-provider-profit rule & 0 & weekly profit & Separates non-negative provider economics from loss-making periods in profitability-based readouts. \\
        \addlinespace
        Sustained-loss window & 3 & weeks & Defines how long negative profitability must persist before profitability-induced churn is treated as a Stage~2 pattern. \\
        \addlinespace
        Liquidity event timing & 20 & week index & Marks the discrete sell event used in the liquidity-shock scenario. \\
        \addlinespace
        Liquidity sell-pressure scalar & 35 & percent & Defines the size of the unlock-style sell event in the liquidity-shock scenario. \\
        \addlinespace
        Competitive-yield scalar & 1.5 & relative multiple & Sets the strength of the outside opportunity used in the competitive-yield scenario. \\
        \addlinespace
        Provider-cost multiplier & 1.35 & multiplier & Sets the exogenous increase in operating costs used in the cost-inflation scenario. \\
        \addlinespace
        Price-signal threshold & 10 & percent & Defines a material price deviation from baseline in Stage~1 detection. \\
        \addlinespace
        Utilization-signal threshold & 10 & percent & Defines a material utilization deviation from baseline in Stage~1 detection. \\
        \addlinespace
        Profitability-signal threshold & 10 & percent & Defines a material profitability deviation from baseline in Stage~1 detection. \\
        \addlinespace
        Provider-count threshold & 5 & percent & Defines a material active-provider deviation from baseline in Stage~1 detection. \\
        \addlinespace
        Churn-increase threshold & 20 & percent & Defines a material churn deviation from baseline in Stage~1 detection. \\
        \bottomrule
    \end{tabularx}
    \caption{Excerpt of the fixed thresholds used in the DTSE reporting logic.}
    \label{tab:app_dtse_threshold_registry_excerpt}
\end{table}

\subsection{Override Ledger}
\label{app:dtse_override_ledger}
The reported DTSE runs also include a small number of profile-specific and global overrides. Table~\ref{tab:app_dtse_override_ledger} summarizes the main ones in human-readable form.

\begin{table}[H]
    \centering
    \small
    \renewcommand{\arraystretch}{1.15}
    \setlength{\tabcolsep}{4pt}
    \begin{tabularx}{\textwidth}{@{}>{\raggedright\arraybackslash}p{0.24\textwidth}>{\raggedright\arraybackslash}p{0.18\textwidth}>{\raggedright\arraybackslash}p{0.16\textwidth}X@{}}
        \toprule
        \textbf{Override Family} & \textbf{Scope} & \textbf{Status} & \textbf{Purpose} \\
        \midrule
        Profile-specific churn thresholds & ONO, Helium, Geodnet & Modeled assumption & Allow continuation sensitivity to differ across mechanism profiles without treating those values as empirical estimates of live behavior. \\
        \addlinespace
        ONO maximum weekly churn cap & ONO & Modeled assumption & Prevents unrealistic discontinuities in weekly exit intensity while preserving participation stress under adverse conditions. \\
        \addlinespace
        Provider-type switching adjustment & Global competitive-yield rule & Scenario logic & Allows competitive-yield pressure to affect provider types differently under the same outside opportunity. \\
        \bottomrule
    \end{tabularx}
    \caption{Main profile-specific and global overrides used in the reported DTSE runs.}
    \label{tab:app_dtse_override_ledger}
\end{table}

\subsection{DTSE Run Manifest}
\label{app:dtse_run_manifest}
Table~\ref{tab:app_dtse_manifest_excerpt} records the key features of the frozen run manifest used for Sections~\ref{sec:simulation_results} and~\ref{sec:discussion}.

\begin{table}[H]
    \centering
    \small
    \renewcommand{\arraystretch}{1.15}
    \setlength{\tabcolsep}{5pt}
    \begin{tabularx}{\textwidth}{@{}>{\raggedright\arraybackslash}p{0.32\textwidth}X@{}}
        \toprule
        \textbf{Manifest Field} & \textbf{Recorded Value} \\
        \midrule
        Run timestamp (UTC) & 24 February 2026, 10:21 UTC \\
        Scenario-grid identifier & DTSE Chapter 7 grid, version 1 \\
        Deterministic seed policy & Master seed plus run index under a fixed seed schedule \\
        Horizon and repetitions & 52 weeks per run; 60 repeated runs per scenario--profile condition \\
        Mechanism profiles & ONO calibrated profile, Helium BME profile, Hivemapper profile, Grass profile, and Geodnet profile \\
        Scenario set & Baseline Neutral, Demand Contraction, Liquidity Shock, Competitive Yield Pressure, and Provider Cost Inflation \\
        \bottomrule
    \end{tabularx}
    \caption{Run-manifest excerpt for the frozen DTSE reporting set.}
    \label{tab:app_dtse_manifest_excerpt}
\end{table}

\paragraph{Empirical applicability note}
The empirical metric-applicability boundary remains in Section~\ref{sec:empirical_stress_layer}, where the thesis defines the \emph{N/R} rule, applies the comparator-level matrix, and states the Onocoy-specific observability limits directly in the main body rather than repeating them in appendix form.

\subsection{Semi-Structured Interview Context Summary}
\label{app:onocoy_interview_context}

The thesis also draws on a semi-structured interview with a representative of the Onocoy project. The purpose of the interview was not to establish mechanism facts in place of public documentation, but to clarify underspecified operational issues and identify stress-relevant sensitivities where public materials remain thinner than the token and payment architecture itself. Accordingly, the interview is treated here as bounded practitioner context. It informs interpretation and evidence-gap identification, but it does not override protocol documentation, public tokenomics materials, or directly auditable public data.

\paragraph{Emission structure}
\emph{Question focus:} Is issuance time-based, demand-based, governance-adjustable, or hybrid? Is there a formal taper plan? \\
\emph{Interview clarification:} Issuance was described as time-based rather than demand-adjusted in the short run. Public tokenomics documentation also specifies a 16\% annual reduction in newly distributed ONO. \\
\emph{Use in this thesis:} This supports the reading that demand shocks primarily weaken buyback strength rather than automatically reducing issuance pressure; in the thesis, this remains bounded practitioner context consistent with public tokenomics documentation.

\paragraph{Demand measurement}
\emph{Question focus:} How is demand measured internally? \\
\emph{Interview clarification:} Data Credits were described as the main internal demand proxy, with service usage observed through credit purchase and burn activity. Public documentation also makes clear that Data Credits are the non-transferable service unit priced against fiat-denominated value. \\
\emph{Use in this thesis:} This reinforces the thesis treatment of Data Credits as the clearest available demand-linked field in the Onocoy design; it is used as bounded practitioner context broadly aligned with public documentation.

\paragraph{Buyback logic}
\emph{Question focus:} What is the current burn-to-emission relationship? How does demand affect net supply in the short run? \\
\emph{Interview clarification:} Buybacks were described as revenue-mediated rather than as a strict mechanical burn-and-mint equilibrium. Public tokenomics documentation likewise ties Data Credit usage to burn and conditional ONO buyback routing rather than to a native BME rule. \\
\emph{Use in this thesis:} This supports the thesis distinction between ONO and native BME systems and clarifies why demand depth matters for token-side support; it is treated as bounded practitioner context consistent with public tokenomics routing language.

\paragraph{Provider economics}
\emph{Question focus:} What kinds of provider-cost differences and economic horizons matter in practice? \\
\emph{Interview clarification:} Provider economics were described as heterogeneous rather than uniform, with hardware choice, reward levels, and token price all affecting economic viability. Public retail listings likewise suggest that Onocoy-compatible hardware can span a few-hundred-dollar entry level, while more complete setups are materially higher in cost. \\
\emph{Use in this thesis:} This is used only to justify why provider-cost sensitivity should remain explicit in interpretation and DTSE assumption design; it combines bounded interview clarification with dated public market-price context rather than a fixed mechanism fact.

\paragraph{Retention sensitivity}
\emph{Question focus:} Which mechanism most strongly influences provider retention today? \\
\emph{Interview clarification:} Retention was described as more sensitive to profitability, macro conditions, and local saturation than to governance participation. \\
\emph{Use in this thesis:} This supports the thesis view that participation elasticity may emerge through provider economics before broader service deterioration becomes obvious; it remains bounded practitioner context, not a documented mechanism rule.

\paragraph{Governance response}
\emph{Question focus:} What happens operationally if token price drops sharply, and who can change emissions or related rules? \\
\emph{Interview clarification:} Stress adaptation was described as discretionary rather than automatic, with rule changes depending on explicit governance action rather than built-in feedback loops. \\
\emph{Use in this thesis:} This supports the thesis distinction between algorithmic stabilization and governance-mediated intervention; it is treated as bounded practitioner context consistent with the current public evidence boundary.

\paragraph{Utility versus speculation}
\emph{Question focus:} Do participants appear mainly utility-driven, speculative, or mixed in their token behavior? \\
\emph{Interview clarification:} Participant behavior was described as mixed: some operators appear long-horizon, while others sell rewards to offset operating costs. \\
\emph{Use in this thesis:} This supports the thesis reading that price volatility can amplify margin stress without reducing the network to a purely speculative system; it remains bounded practitioner context only.
