% SCOPE CONTRACT: EMPIRICAL ANALYSIS (Section 6)
% Content derived from Empirical Resilience Research (Helium, Geodnet, Hivemapper)

\section{Empirical Stress Analysis}
\label{sec:empirical_analysis}

\subsection{Introduction}
Decentralized Physical Infrastructure Networks (DePIN) use cryptographic incentives to coordinate real-world hardware across sectors such as wireless connectivity and geospatial services \cite{FrontiersDePIN2025}. Unlike purely digital crypto systems, DePIN networks operate under physical constraints such as hardware sunk costs, geographic friction, and operational latency, which materially change risk transmission \cite{Ballandies2023, Lin2024}. Sustainability therefore depends on the interaction between mechanism design and provider behavior under stress \cite{Chiu2024}.

This chapter provides a retrospective empirical layer that complements DTSE. ``Retrospective'' means it evaluates documented historical stress windows in live networks; it does not run counterfactual scenarios. ``Empirical'' means it reports observed stress behavior and comparable diagnostics, without treating those observations as proof of future outcomes.

Historically, cryptoeconomic evaluation has relied heavily on ABM and stochastic simulation \cite{Morris2019, Collins2024}. Those tools remain essential, but outputs are assumption-conditional. Empirical event-window analysis adds a second evidence layer by showing how comparable systems behaved in realized stress windows \cite{MacKinlay1997}.

Accordingly, this chapter uses an event-study-based static stress framework. It benchmarks Onocoy against Helium, Hivemapper, Render, and Geodnet under observed liquidity, demand, and competitive-yield pressure windows. The purpose is diagnostic: identify observed failure signatures and map them to the metric vocabulary later used in DTSE.

\subsection{Theoretical Framework: Operationalizing Stress Without Simulation}
To analyze DePIN robustness in an empirical layer, this chapter combines comparative statics with empirical event studies \cite{MacKinlay1997}. Here, comparative statics means comparing observable metrics across defined regimes, and event study means measuring deviations around specific historical shocks. Stress is therefore defined as an observed market-and-network regime, not as a simulated perturbation.

\subsubsection{Defining Stress Factors in Physical Networks}
In this framework, stress factors are exogenous shocks that challenge economic viability. Three stress families are used \cite{Lin2024, Ho2022}:

\begin{itemize}
    \item \textbf{The Subsidy Gap (Reward Addiction):} the gap between provider CapEx/OpEx and usage-linked fiat value entering the network. In early phases this gap is often bridged by token incentives; when token value compresses, the gap can widen \cite{Lin2024, McConaghy2020}. In DePIN terms, this asks whether usage-linked value can fund participation without persistent subsidy.
    \item \textbf{Speculative Fragility:} the sensitivity of participation proxies (uptime/retention/activity) to token-price volatility. High fragility means participation degrades quickly when token-denominated rewards lose fiat value \cite{Ho2022}. In DePIN terms, this asks how tightly hardware participation tracks token price stress.
    \item \textbf{Competitive Yield Pressure:} participation elasticity when comparable external opportunities offer higher returns for similar hardware profiles \cite{Lin2024, Chiu2024}. In DePIN terms, this captures whether operators can reallocate capacity when adjacent networks offer better reward conditions.
\end{itemize}

\subsubsection{Mechanism Design as a Static Defense}
Within a static framework, mechanism design is evaluated by its ability to absorb stress:

\begin{itemize}
    \item \textbf{Burn-and-Mint Equilibrium (BME):} in common DePIN implementations, users burn token-linked credits while provider rewards are distributed via issuance channels \cite{RenderBME}. A key static check is whether usage-linked burn can keep pace with reward outflow in adverse windows.
    \item \textbf{Capped Supply with Programmed Distribution:} for Onocoy, publicly documented mechanics indicate a capped/pre-minted supply with rule-based distribution and buyback-burn routing, rather than uncapped new creation \cite{OnocoyWhitepaper301, OnocoyToken, OnocoyRewards, OnocoyBonus}. The static check is whether distribution policy and usage-linked value can sustain participation as reward intensity normalizes.
    \item \textbf{Sunk Cost Moats:} physical hardware is often illiquid and location-bound, creating real exit friction that can delay churn during drawdowns \cite{Arthur1989, Ballandies2023}. In DePIN terms, this is a retention buffer from physical frictions, not necessarily from incentive strength.
\end{itemize}

\subsubsection{Methodology: Empirical Event Studies}
This chapter applies event-study logic from financial econometrics \cite{MacKinlay1997}. Historical shocks (for example, broad crypto drawdowns, mechanism transitions, and reward-structure changes) are treated as event windows. Within those windows, participation and incentive metrics are compared against pre-event baselines to identify deviations in churn, utilization, and solvency proxies.

This empirical layer does not estimate causal treatment effects. It provides structured comparative evidence for stress signatures that later inform DTSE scenario design and interpretation boundaries.

\subsection{Comparative Scope: The Solana DePIN Ecosystem}
This chapter uses a Solana-centric comparator set to keep mechanism and data conventions reasonably aligned. Existing DePIN ecosystem analyses describe Solana as a major execution layer in current market structure \cite{FrontiersDePIN2025, Messari2024}; this thesis uses that context as a practical sampling frame, not as a normative claim.

\subsubsection{The Comparative Set}
Projects are grouped by hardware and operating-cost profile, because those factors strongly influence stress transmission:

\subsubsection*{Archetype A: Commodity Sensor Networks (Low OpEx, Moderate CapEx)}
\begin{itemize}
    \item \textbf{Helium (IoT/Mobile):} mature hotspot network used here as a retention and migration-stress reference case \cite{FrontiersDePIN2025}.
    \item \textbf{Geodnet (GNSS/RTK):} adjacent GNSS network with explicit location-governance constructs (\emph{Location NFT}) relevant to supply-density and yield management \cite{GeodnetLocationNFT}.
    \item \textbf{Onocoy (GNSS/RTK):} focal case in this thesis; capped/pre-minted supply with documented distribution and reward program rules \cite{OnocoyWhitepaper301, OnocoyToken, OnocoyRewards}.
\end{itemize}

\subsubsection*{Archetype B: High-Performance Compute (High OpEx, High CapEx)}
\begin{itemize}
    \item \textbf{Render:} GPU-intensive provider base used as a compute-oriented comparator for participation sensitivity under cost and yield pressure \cite{Lin2024, FrontiersDePIN2025}.
    \item \textbf{io.net:} aggregated compute network included as a compute-market comparator in this chapter's sample.
\end{itemize}

\subsubsection*{Archetype C: Active-Labor Physical Work}
\begin{itemize}
    \item \textbf{Hivemapper:} mapping network requiring ongoing active labor; participation is therefore sensitive to variable effort-cost coverage \cite{Chiu2024, FrontiersDePIN2025}.
\end{itemize}

\subsubsection{Comparator Lineage and Scope Funnel}
To keep scope transitions explicit, this chapter separates three layers: the empirical universe used for retrospective comparison in this chapter, the event-window sample that remains analytically emphasized, and the frozen DTSE profile set used in Chapter~\ref{sec:simulation_results}. This avoids scope ambiguity when readers compare Chapter~\ref{sec:empirical_analysis} with the DTSE manifest.

\begin{table}[ht]
    \centering
    \footnotesize
    \renewcommand{\arraystretch}{1.2}
    \setlength{\tabcolsep}{4pt}
    \begin{tabularx}{\textwidth}{@{}>{\raggedright\arraybackslash}p{0.20\textwidth}>{\raggedright\arraybackslash}p{0.30\textwidth}>{\raggedright\arraybackslash}p{0.22\textwidth}X@{}}
        \toprule
        \textbf{Comparison Layer} & \textbf{Set Used} & \textbf{Primary Role} & \textbf{Boundary / Reason} \\
        \midrule
        Chapter~\ref{sec:empirical_analysis} comparator universe &
        Helium, Geodnet, Onocoy, Render, Hivemapper, io.net &
        Retrospective empirical stress benchmarking under observed windows &
        Uses publicly documented projects with heterogeneous hardware profiles; metric applicability gates (Section~\ref{subsec:metric_applicability_framework}) determine which comparisons are inferentially valid. \\
        \addlinespace
        Chapter~\ref{sec:empirical_analysis} stress-window emphasis &
        Helium, Geodnet, Onocoy, Hivemapper (plus contextual references to Render/io.net where relevant) &
        Channel-specific interpretation of observed stress signatures &
        Emphasis follows event-window observability and metric applicability constraints (including N/R policy). \\
        \addlinespace
        Chapter~\ref{sec:simulation_results} frozen DTSE profile set &
        \path{ono_v3_calibrated, helium_bme_v1, hivemapper_v1, grass_v1, geodnet_v1} &
        Controlled counterfactual comparison under matched DTSE stress inputs &
        Fixed by frozen manifest (Appendix Section~\ref{app:dtse_run_manifest}); profile inclusion reflects implementation-ready DTSE rule mappings rather than the broader empirical universe. \\
        \bottomrule
    \end{tabularx}
    \caption{Comparator lineage from empirical scope to frozen DTSE profile set.}
    \label{tab:comparator_lineage_funnel}
\end{table}

\subsubsection{Metric Standardization for Static Comparison}
To compare heterogeneous systems consistently, this chapter uses standardized metrics and explicit formulas:

\begin{table}[htbp]
    \centering
    \small
    \renewcommand{\arraystretch}{1.35}
    \begin{tabular}{p{0.20\linewidth} p{0.22\linewidth} p{0.46\linewidth}}
        \toprule
        \textbf{Category} & \textbf{Standardized Metric} & \textbf{Definition \& Proxy Utility} \\
        \midrule
        Incentive Solvency & Burn-to-Mint Ratio & \( \mathrm{BMR}_t = \frac{B_t}{M_t} \). Under this chapter's convention, \( \mathrm{BMR}<1 \) indicates subsidy dependence; \( \mathrm{BMR}\ge1 \) indicates burn coverage of reward outflow. \\
        \midrule
        Provider Retention & 30-Day Node Retention & \( \mathrm{Ret}_{30} = \frac{N_{t_0+30}}{N_{t_0}} \), where \(t_0\) is event-window baseline. Captures short-horizon participation persistence under stress. \\
        \midrule
        Economic Efficiency & Revenue per Node & \( \mathrm{RPN}_t = \frac{R_t}{N_t} \). Proxy for average provider-level economic support from observed network revenue. \\
        \midrule
        Speculative Velocity & Token Turnover & \( \mathrm{Turnover}_t = \frac{V_t}{MC_t} \). High values during stress can indicate short-horizon capital rotation dominance \cite{CFA2018}. \\
        \midrule
        Valuation Risk & FDV / Annualized Revenue & \( \mathrm{FDVRev}_t = \frac{\mathrm{FDV}_t}{12 \cdot R^{(m)}_t} \), where \(R^{(m)}_t\) is monthly revenue. Proxy for valuation premium versus current revenue scale. \\
        \bottomrule
    \end{tabular}
    \caption{Standardized metrics for static DePIN comparison.}
    \label{tab:comparative_metrics}
\end{table}

\subsubsection*{Empirical observability and non-imputation}
The standardized metric catalog in Table~\ref{tab:comparative_metrics} provides a common comparative vocabulary for heterogeneous DePIN profiles, but empirical observability differs across mechanisms, maturity stages, and available public data in the analyzed event windows. A metric is reported only when its required inputs are publicly observable at the necessary resolution and can be computed under consistent definitions without imputing missing values. Where these conditions are not met, the metric is marked \emph{N/R} (not reliably observable) and accompanied by a short reason code. Metrics marked \emph{N/R} are not used for comparative inference in this empirical chapter. This reporting rule applies to the empirical layer only and does not alter DTSE computation in later chapters, where the pre-specified metric set is produced consistently under the simulator's controlled assumptions.

\noindent\textbf{N/R reason codes:}
\begin{itemize}
    \item \textbf{NR1} Required inputs are not publicly observable at the necessary resolution for the event window.
    \item \textbf{NR2} Native metric definitions differ materially across protocols and cannot be harmonized without strong additional assumptions.
    \item \textbf{NR3} Market/revenue observability is too thin or distorted in the window for stable liquidity-, valuation-, or revenue-dependent ratios.
\end{itemize}

\subsubsection{Metric Applicability Framework}
\label{subsec:metric_applicability_framework}
To prevent category errors in cross-network comparison, this chapter applies four explicit gates before using any metric for inference.
\begin{enumerate}
    \item \textbf{Mechanism-compatibility gate:} the metric must have a valid interpretation under the underlying mechanism class (for example, burn-to-mint is native in BME designs and adapted for capped-supply reward-distribution designs).
    \item \textbf{Observability gate:} required inputs must be publicly observable at sufficient resolution without imputing missing values (NR1).
    \item \textbf{Harmonization gate:} metric definitions must be alignable across protocols without introducing strong additional assumptions (NR2).
    \item \textbf{Maturity gate:} market/revenue-dependent metrics require sufficient liquidity and revenue depth to avoid unstable denominators and microstructure-dominated noise (NR3).
\end{enumerate}

\noindent\textbf{Comparative inference rule:} a metric marked \emph{N/R} for a given project-window is excluded from comparative ranking in this empirical chapter. It may still appear as context, but it is not used as evidence for cross-project ordering.

\begin{table}[ht]
    \centering
    \small
    \renewcommand{\arraystretch}{1.2}
    \setlength{\tabcolsep}{4pt}
    \begin{tabularx}{\textwidth}{@{}>{\raggedright\arraybackslash}p{0.24\textwidth}>{\centering\arraybackslash}p{0.05\textwidth}>{\centering\arraybackslash}p{0.05\textwidth}>{\centering\arraybackslash}p{0.05\textwidth}>{\centering\arraybackslash}p{0.05\textwidth}>{\centering\arraybackslash}p{0.06\textwidth}X@{}}
        \toprule
        \textbf{Metric} & \textbf{He} & \textbf{Ge} & \textbf{On} & \textbf{Re} & \textbf{Hi} & \textbf{Primary Constraint} \\
        \midrule
        Burn-to-Mint Ratio & A & A & P & A & A & Onocoy uses an adapted capped-supply burn-to-distribution analogue. \\
        30-Day Node Retention & A & A & P & A & A & Onocoy stress-window history remains short; inference is directional. \\
        Revenue per Node & A & A & N/R & A & P & Onocoy revenue depth is immature (NR3); Hivemapper has labor-intensity comparability caveats. \\
        Token Turnover & A & A & N/R & A & P & Onocoy post-TGE liquidity depth is insufficient for stable inference (NR3). \\
        FDV / Annualized Revenue & A & A & N/R & A & P & Onocoy denominator instability under early-stage revenue conditions (NR3). \\
        \bottomrule
    \end{tabularx}
    \caption{Metric applicability matrix for empirical comparison windows (He = Helium, Ge = Geodnet, On = Onocoy, Re = Render, Hi = Hivemapper; A = applicable; P = partially applicable; N/R = not reliably observable).}
    \label{tab:metric_applicability_matrix}
\end{table}

\subsubsection*{Stress-channel metric emphasis}
Table~\ref{tab:stress_metric_applicability_map} makes the scenario-to-metric logic explicit for empirical interpretation. This map is used to choose the first-signal metric family in each empirical stress channel and to avoid over-weighting low-signal metrics in unsuitable windows.

\begin{table}[ht]
    \centering
    \small
    \renewcommand{\arraystretch}{1.2}
    \setlength{\tabcolsep}{4pt}
    \begin{tabularx}{\textwidth}{@{}>{\raggedright\arraybackslash}p{0.20\textwidth}>{\raggedright\arraybackslash}p{0.24\textwidth}>{\raggedright\arraybackslash}p{0.18\textwidth}X@{}}
        \toprule
        \textbf{Stress Channel} & \textbf{Most Informative Metric Families} & \textbf{Lower-Signal Metrics} & \textbf{Interpretive Note} \\
        \midrule
        Liquidity shock & Token turnover; retention/churn & Revenue per node; FDV/revenue & Price-liquidity dislocation is expected to register first in market-participation channels. \\
        Competitive yield pressure & Retention/churn; revenue per node (where available) & Token turnover; FDV/revenue & Provider reallocation can occur without a simultaneous market-wide price event. \\
        Demand contraction & Burn-to-mint; utilization-linked measures; revenue per node (where available) & Token turnover (initially) & Usage-linked channels typically move before participation channels in this stress family. \\
        Provider-cost pressure & Provider profitability; retention/churn & Token turnover; FDV/revenue & Cost-side stress is expected to register first in economics fields and then in participation fields. \\
        \bottomrule
    \end{tabularx}
    \caption{Stress-channel to metric applicability map for empirical interpretation.}
    \label{tab:stress_metric_applicability_map}
\end{table}

\subsubsection*{Onocoy applicability profile}
\label{subsec:onocoy_metric_applicability_profile}
For this empirical chapter, Onocoy is not scored as if all five standardized metrics were equally informative. The current applicability profile is:
\begin{itemize}
    \item \textbf{Primary signals now:} retention/churn and adapted burn-to-mint trends.
    \item \textbf{Directional-only signals:} adapted burn-to-mint and short-window retention, both interpreted with early-stage caveats.
    \item \textbf{N/R in this chapter:} revenue per node, token turnover, and FDV/annualized revenue under current observability and maturity conditions (NR3).
\end{itemize}
This chapter therefore uses an ``applicable metrics only'' inference policy for Onocoy. Full applicability notes and temporal-usage conditions are catalogued in Appendix Section~\ref{app:empirical_metric_applicability}.

\subsection{Empirical Stress Test: Historical Performance Analysis}
This section applies the static framework to documented stress windows. The goal is descriptive benchmarking of observed behavior, not causal proof or forecast generation.

\subsubsection{Stress Scenario 1: Liquidity Shock and Crypto Winter (2022--2023)}
\textbf{Context:} Exchange primary market data for HNT/USDT show severe price compression during the 2022--2023 window. This period is used as an empirical reference for liquidity-driven incentive compression under broad adverse conditions \cite{BinanceHNTUSDTWeekly2022_2023,Ho2022}.

\textbf{Observed pattern (reported):}
\begin{itemize}
    \item Active hotspot counts did not collapse one-for-one with price drawdown in reported windows, consistent with physical exit friction and sunk-cost effects \cite{Ballandies2023}.
    \item Growth and maintenance dynamics appeared weaker than aggregate installed-base persistence, consistent with latent capacity-quality erosion risk.
    \item Governance-level structural response (migration/sub-DAO architecture) was proposed to reconfigure incentive and governance surfaces under stress \cite{HeliumHIP70}.
\end{itemize}

\textbf{Implication for Onocoy:} Installed GNSS infrastructure may remain comparatively sticky during drawdowns due to deployment friction, while new growth and maintenance intensity may degrade earlier. This chapter treats that as motivation for DTSE sensitivity testing, not as protocol-level projection.

\subsubsection{Stress Scenario 2: Competitive Yield Pressure (2024--2025)}
\textbf{Context:} As DePIN verticals mature, adjacent networks increasingly compete for overlapping provider profiles and installation geographies \cite{Lin2024, Chiu2024}.

\textbf{Observed pattern (reported):}
\begin{itemize}
    \item GNSS-like settings show explicit density and location-governance controls (for example, Location NFT logic) used to shape local yield conditions \cite{GeodnetLocationNFT}.
    \item Where multi-homing is feasible, operators can rebalance toward higher near-term yield or stronger liquidity conditions.
    \item Reported allocation behavior is consistent with elastic provider participation under external yield pressure, rather than purely demand-driven participation changes.
\end{itemize}

\textbf{Implication for Onocoy:} Competitive yield pressure is a first-order risk channel for GNSS networks when compatible hardware can support more than one ecosystem. In this chapter, that pattern is treated as an empirical input for later DTSE competitive-yield scenarios.

\subsubsection{Stress Scenario 3: Operational Cost Pressure (Hivemapper 2024)}
\textbf{Context:} Hivemapper participation requires ongoing active labor, so effective contribution cost includes time and fuel as variable OpEx components \cite{Chiu2024, FrontiersDePIN2025}.

\textbf{Observed pattern (reported):}
\begin{itemize}
    \item When token-denominated reward value weakens, participation can plateau or decline faster than in passive, low-maintenance sensor setups.
    \item Reported mitigation efforts combined demand-side enterprise routing with targeted incentive programs, consistent with incentive re-targeting responses.
\end{itemize}

\textbf{Implication for Onocoy:} This case highlights cost-floor sensitivity as a general stress channel: once real operating effort exceeds expected reward value, participation elasticity increases.

\subsubsection*{Historical Stress Windows at a Glance}
\begin{table}[ht]
    \centering
    \small
    \renewcommand{\arraystretch}{1.2}
    \begin{tabularx}{\textwidth}{@{}p{0.26\linewidth}p{0.34\linewidth}X@{}}
        \toprule
        \textbf{Window} & \textbf{Primary Observed Signal (Reported)} & \textbf{DTSE Link Used in This Thesis} \\
        \midrule
        Liquidity shock and crypto winter (2022--2023) & Price compression with weaker growth/maintenance dynamics despite installed-base persistence. & Motivates liquidity-shock channel and separation between immediate price stress and delayed participation effects. \\
        Competitive yield pressure (2024--2025) & Provider reallocation behavior under adjacent-network yield differentials where multi-homing is feasible. & Motivates competitive-yield scenario and early participation-sensitivity diagnostics. \\
        Operational cost pressure (Hivemapper 2024) & Participation sensitivity to variable effort/OpEx when reward value weakens. & Motivates provider-cost inflation channel and profitability-first stress sequencing. \\
        \bottomrule
    \end{tabularx}
    \caption{Historical stress windows summarized for transition into DTSE scenario design.}
    \label{tab:empirical_stress_windows_summary}
\end{table}

\subsection{Tokenomic Mechanism Analysis: Comparative Robustness}
Using the preceding empirical benchmarks, this section interprets Onocoy's mechanism position relative to comparator designs under stress.

\subsubsection{Burn-and-Mint Equilibrium (BME)}
Several comparator DePIN networks use BME-like value-accrual pathways in which user-side demand burns token-linked credits while provider rewards are distributed via issuance channels \cite{RenderBME}.
\begin{itemize}
    \item In stronger demand regimes, burn can offset reward outflow.
    \item In weak demand regimes, issuance pressure can dominate burn, weakening incentive-solvency proxies \cite{FrontiersDePIN2025, Lin2024}.
\end{itemize}

\textbf{Onocoy variance:} Onocoy documentation describes capped/pre-minted supply with programmed distribution, plus buyback-and-burn routing from revenue channels \cite{OnocoyToken, OnocoyRewards, OnocoyBonus}.

\textbf{Interpretive tradeoff:} Relative to flexible issuance-heavy designs, capped/pre-minted distribution can reduce inflationary tail risk but may offer less short-run subsidy flexibility if reward value and participation economics diverge materially.

\subsubsection{Incentive Solvency: Burn-to-Mint Ratio}
For this chapter, burn-to-mint is the principal solvency proxy:
\[
\mathrm{BMR}_t=\frac{B_t}{M_t}.
\]
Under this reporting convention, ``mint'' denotes reward outflow/distribution as measured in the metric definition; for Onocoy this is interpreted as distribution from fixed inventory pathways rather than uncapped token creation.

\begin{itemize}
    \item Reported mature windows can approach parity (\(\mathrm{BMR}\approx1\)) in some regimes.
    \item Early-stage systems often remain below parity while demand-side channels scale.
    \item Persistent sub-unity behavior indicates continuing subsidy dependence under this chapter's metric convention.
\end{itemize}

\subsection{DePIN Failure Modes: Diagnostic Matrix}
Based on the empirical cases and the stress framework, this chapter uses five operational failure-mode labels as diagnostic categories (not projections) \cite{Braakman2022, Morris2019}.

\begin{enumerate}
    \item \textbf{Reward--Demand Decoupling:} reward outflow remains high while usage-linked sink strength weakens \cite{FrontiersDePIN2025, OnocoyRewards}. In DePIN terms, rewards continue while service demand lags.
    \item \textbf{Liquidity-Driven Incentive Compression:} market-side token dislocation reduces fiat-equivalent reward value and compresses provider economics \cite{Ho2022, Lin2024}. In DePIN terms, effective returns fall even when physical costs are stable.
    \item \textbf{Latent Capacity Degradation:} participation appears stable while maintenance intensity and quality buffers erode \cite{Arthur1989, Ballandies2023}. In DePIN terms, reported node counts can mask declining resilience.
    \item \textbf{Elastic Provider Exit:} operators reallocate to better external yields when hardware compatibility permits \cite{Chiu2024, Lin2024, GeodnetLocationNFT}. In DePIN terms, supply can move faster than demand.
    \item \textbf{Dilution Feedback Loop:} weaker token value reduces effective reward, participation, and service confidence, which can further pressure value channels \cite{Lin2024, FrontiersDePIN2025}. In DePIN terms, stress channels can reinforce each other.
\end{enumerate}

\subsection{Conclusion: Benchmarking Summary}
This chapter shows that empirical, retrospective stress analysis is a useful complement to DTSE. Comparative event windows suggest recurring stress signatures across hardware-intensive tokenized networks: physical sunk costs can buffer immediate churn, but prolonged subsidy dependence and weak usage-to-reward coupling are associated with higher fragility risk.

For Onocoy, two dimensions are carried forward into DTSE: (i) reward-demand coupling strength and (ii) provider-participation sensitivity under external yield and cost pressure.

\subsection{Bridge to Simulation}
Historical evidence identifies signatures in realized windows, but it cannot test unobserved thresholds under controlled counterfactual conditions for a new capped/pre-minted profile. This gap motivates DTSE in subsequent chapters: simulation is used as an evaluative instrument under explicit assumptions, not as a projection tool.
