% Provisional source-preserving import from Person A draft snapshot.
% Source: tmp/docs/personA_submission_extracted.txt (P0156-P0181)

The following imported block preserves Person A's interview-integration framing for later harmonization with the methodology chapter.

Person A's draft introduces a semi-structured interview integration layer intended to complement theory and public documentation with practitioner-level context. The interviews are treated as support for identifying assumptions and operational frictions that may not be fully visible in public protocol material.
% ARCHIVE-CITE-NOTE: [Person A sources: Doring and Bortz 2015; Messari 2023]

The draft explicitly treats interview signals as indicative rather than deterministic. In this framing, statements regarding provider payback expectations, churn drivers, internal demand indicators, and emission-adjustment thinking are interpreted as calibration ranges for modeling, rather than as direct empirical facts.
% ARCHIVE-CITE-NOTE: [Person A sources: Mayring and Fenzl 2022]

A disclosure constraint is also preserved from the submission: no confidential internal data is published in this thesis section, and interview-derived inputs are incorporated as bounded assumptions.

Person A's limitations framing is preserved in condensed form:
\begin{itemize}
    \item granular provider-level financial data is not fully public,
    \item token-market behavior can diverge from stylized equilibrium assumptions,
    \item interview responses are subjective and context-bound,
    \item descriptive sections should not be interpreted as causal inference.
\end{itemize}

This block is kept provisional so that a later pass can reconcile terminology and references with the final DTSE methodology narrative.

\paragraph{Traceability note.} Full verbatim Person A submission snapshots are preserved in Appendix~\ref{subsec:contributor_draft_archive} for audit and attribution during drafting.
